---
title: "Political Beliefs, Views about Technocracy, and Energy and Climate Policy Preferences"
bibliography: /Users/nowlinmc/Dropbox/refs.bib
csl: /Users/nowlinmc/Dropbox/Projects/Manuscripts/csl/sage-harvard.csl
output:
  word_document: default
  pdf_document:
    keep_tex: yes
    template: /Users/nowlinmc/Dropbox/Projects/Manuscripts/article.latex
...

# Introduction

Addressing the rising energy and environmental challenges of the 21st century will require a high degree of scientific and technological knowledge and expertise. However, the role of science in democratic decision-making is fraught with tension between the merits of expertise and the importance of values and representation in policymaking. Values inform thinking about the role of government, what is considered as a risk, and the types of behavior that society should regulate, among other considerations. The tensions between expert and democratic decision-making are prominent in energy and environmental issues where science and values may lead to opposing answers to questions of policy.  

Generally speaking, the US public has a high degree of confidence in science and its ability to improve societal conditions. According to a recent survey from the _Pew Research Center_, 73 percent of respondents believed that science has had a mostly positive impact on society and 86 percent had a great deal or a fair amount of confidence that scientists act in the public interest.^[See @funkKeyFindingsAmericans2020, available here: https://www.pewresearch.org/fact-tank/2020/02/12/key-findings-about-americans-confidence-in-science-and-their-views-on-scientists-role-in-society/] Additionally, the Pew survey found that, overall, 60 percent of those in the US believe that scientists should play an active role in policy debates. However, there are stark partisan differences regarding the influence of scientists in policymaking with 73 percent of Democrats agreeing that scientists should play a role, whereas, 43 percent of Republicans agreed. The results from Pew suggest that the US public is largely supportive of the scientific enterprise, its impacts on society, and its role in informing policy decisions, yet there is a high degree of political polarization regarding the policy influence of science. In this paper, I examine public support for technocracy -- science and expertise driven decision-making -- as a way to make decisions regarding energy and environmental issues as well as the impact of support for technocratic decision-making on views about specific, politically controversial energy sources and climate change policies.^[Throughout the paper I use the terms technocracy and technocratic decision-making interchangeably.] 

While there is a large literature on political beliefs and views about science and science-based political issues [@blankDoesPartisanshipShape2015; @ellisonConservativeProtestantismPublic1995; @gauchatPoliticizationSciencePublic2012; @jelenReligionPartisanshipAttitudes2014; @lachapellePublicPerceptionsExpert2014; @mccrightInfluencePoliticalIdeology2013; @nisbetPartisanBrainHow2015; @rutjensNotAllSkepticism2018; @washburnScienceDenialPolitical2018] as well as a growing literature on anti-intellectualism among the public [@merkleyAntiIntellectualismPopulismMotivated2020; @mottaDynamicsPoliticalImplications2018], there is less work addressing how the public views experts and or technocratic decision-making [as noted by @dommettWhatWeKnow2019]. Here, I develop a three question technocratic scale and use it to examine the influence of political beliefs on views about technocracy. Next, I explore how support for technocratic decision-making informs energy and environmental views and policy preferences. Specifically, I use the technocratic scale to predict support for four controversial energy sources -- nuclear; oil and gas pipelines; hydraulic fracturing ("fracking"); and offshore drilling -- as well as four climate policy approaches including international agreements, Environmental Protection Agency (EPA) regulations, a carbon tax, and a cap-and-trade program. Finally, I examine the interaction of political beliefs and views about technocracy on support for each energy source and climate policy. 

Overall, I find that support for technocracy is highest among liberal Democrats, moderate/conservative Democrats, and moderate/liberal Republicans as opposed to moderate independents and conservative Republicans. I also find that support of technocracy is associated with increased support for all four sources of energy and all four climate policies. In terms of interactions effects, I find that views about technocracy are most impactful for those in the political "middle" (moderate/conservative Democrats and moderate/liberal Republicans) with regard to energy preferences. However, regarding support for the various climate policies, those at either end of the spectrum -- liberal Democrats and conservative Republicans -- were the most impacted by views about technocracy. While preliminary, these findings point to the public's views about technocracy as an important, and understudied, component of their views about controversial energy and environmental issues.   

<!--  aspect of the use of expert-driven decision-making in liberal democracies. --> 

# Technocratic Decision-Making 

How should science and technical information inform policymaking? In democracies, how should science and other types of expertise be weighed against what the public is believed to want? An inherent tension exists within democracies regarding the role of science and technical expertise in policymaking [@jasanoffFifthBranchScience1990]. On the one hand, a strictly technocratic approach could be used where decisions are made by scientists and other experts and other stakeholders are excluded or their participation is otherwise limited. On the other hand, a pure-democratic approach, where the people are sovereign, would keep decision-making in the hands of the public and/or their elected representatives. 

With a technocratic approach, policy decisions in highly technical areas such as energy and the environment are elite driven and deference is made to scientists and others on the basis of their expertise in these areas. Further, since scientists and the scientific process is seen as objective, decisions are made by applying sound science in a value-neutral way. Therefore, policy decisions are based largely on the best available evidence as understood by the experts. Technocratic decision-making extends into all aspect of policymaking including problem identification and policy implementation [@caramaniWillVsReason2017]. One aspect of technocracy is that positions of authority are gained solely through expertise, which leads to political power shifting toward elite experts [@gunnellTechnocraticImageTheory1982; @burrisTechnocraticOrganizationControl1989]. As a result, a pure technocracy is likely to exclude legitimate voices from the debate, which may result in sub-optimal or inequitable outcomes as well as undermine democratic decision-making. Indeed, technocracy and democracy can be seen as zero-sum; when one increases the other decreases [@gilleyTechnocracyDemocracySpheres2017].     

In a democracy, policy decisions -- even those involving complex and technical issues -- are made by the public and/or those that the public elect. A democratic decision process is pluralistic, with many competing ideas and interests. In a pluralistic democracy science exists as fodder for political debates, and as a result science may become politicized leading to it being downplayed or ignored. Finally, democratic decisions are built on a foundation of public support, yet the public may be poorly informed or inconsistent in their views, which can lead to policy responses that are haphazard at best.  

In practice, the US has elements of both technocracy and democracy. During the 1960s and 1970s the federal government, through the “Great Broadening” [@jonesGreatBroadeningHow2019], expanded the scope of issues with which it was engaged such as education, social welfare, and the environment. As the scope of government grew so too did the need for expertise in those areas. Since the Great Broadening, scientists and technical experts have played an increasing role in informing decision-makers though benefit-cost analysis, risks assessments, and advisory panels, among other avenues [@bozemanWhenScienceUsed2019; @jasanoffFifthBranchScience1990; @raynerDemocracyAgeAssessment2003; @powellScienceEPAInformation1999]. Yet, even as the role of science and expertise grew, that growth was tempered somewhat by a recognition of the importance of public participation in the policymaking process. For example, the Administrative Procedure Act (APA) of 1946 requires a period of public comment during executive branch agency rule-making, and the Clean Air Act of 1970 requires science-based pollution standards coupled with opportunities for public input [@fineTechnocracyDemocracyConflicts2004]. Additionally, as noted by @lahsenTechnocracyDemocracyClimate2005, "In the United States, expert deliberation on scientific evidence is not as shielded from public debate, and control of the production and mobilization of knowledge is relatively looser, allowing greater proliferation of competing risk constructions in government and broader society" (142). As a result, seemingly technocratic decision-making processes become fused with political considerations.    

Despite the importance of technocratic decision-making and its inherent tension with a democratic ethos, little work has directly examined how the public views technocracy specifically. One study by @silvaReconcilingScientistsBeliefs2007 developed a _technocratic disposition_ measure that used three questions to create a seven-point scale, where a value closer to 7 was associated with increased support for technocratic decision-making.^[The questions asked: "1 = Public should decide about advanced technologies, 7 = Experts; Better informed should have more influence (7 = Agree strongly); Even if public uninformed, rely on popular opinion (7 = Disagree strongly)" [@silvaReconcilingScientistsBeliefs2007, 764].] The technocratic disposition scale was administered through a survey given to a sample of scientists associated with the various national laboratories in the US as well as scientists from the Union of Concerned Scientists (UCS). The mean on the seven-point scale was 5.38 for the national lab scientists and 5.01 for the UCS scientists, indicating a general support for technocracy among scientists. @silvaReconcilingScientistsBeliefs2007 go on to examine differences in what scientists believed regarding the risks associated with exposure to radiation versus how those risks should be regulated, and they found that the values and beliefs of scientists (e.g., egalitarian worldview) was associated with a "precautionary shift", where scientists favored regulation beyond what they felt was the actual risk. However, those with an increased technocratic disposition were less likely to demonstrate a precautionary shift. This finding suggests that views regarding technocracy can have an impact on views regarding specific energy issues.    

A second study by @bertsouTechnocraticAttitudesCitizens2017 examined public opinion regarding technocracy using the European Values Survey (EVS) to examine the attitudes of respondents across 27 European Union (EU) countries. Their measure of technocracy was taken from the statement, 

> I’m going to describe various types of political systems and ask what you think about each as a way of governing this country. For each one, would you say it is a very good, fairly good, fairly bad or very bad way of governing this country? _Having experts, not government, make decisions according to what they think is best for the country_    

@bertsouTechnocraticAttitudesCitizens2017 found that support for technocracy _decreases_ as support for democracy _increases_, which points to the public viewing technocracy and democracy as a trade-off. Additionally, they found that increasing trust in political institutions as well as trust in the EU was associated with increased support for technocracy. Finally, @bertsouTechnocraticAttitudesCitizens2017 found that right-wing political orientations are associated with increased support for technocracy, and they speculate that this because technocratic decision-making is associated with free-market economics in the EU. This finding points to the role that political beliefs may play in the formation of attitudes about technocracy. 

While little work has examined how the public views questions of technocracy vs. democracy, there is a related literature on how the public views science and scientists in the policymaking process that could provide insights into questions of support for technocratic decision-making. Technocracy is similar to what @gauchatPoliticalContextScience2015 termed the science-policy cultural image, which "refers to the claim that scientific evidence can improve government policy through the objective evaluation of various policy positions and the identification of social problems" (724). As noted, a seeming contradiction exists in the US, where on the one hand science is highly valued and respected, yet on the other hand in many contentious and polarized policy issues expertise is downplayed by political elites in favor of political and ideological considerations. As a result of elite-driven debate, the public comes to doubt the scientific basis of many of these issues [@hmielowskiAttackScienceMedia2014; @krosnickImpactFall19972000; @merkleyPartyElitesManufactured2018; @gustafsonDevelopmentPartisanPolarization2019], which could possibly lead to a general doubt about technocracy.  

For the US public, polarized scientific or technical issues and the associated political debates has likely solidified a connection between political beliefs and views about science generally, as well as science applied to policy problems. As noted by @gauchatPoliticizationSciencePublic2012, there has been a general decline of trust in science among conservatives over the last few decades. Additional work has found that increasing conservatism is associated with decreased trust in scientists as sources of information about the environment [@brewerWhoseScienceYou2013], trust in science produced by federal agencies is generally lower among conservatives [@myersPredictorsTrustGeneral2017], and Democrats were more likely than those with other party attachments to trust scientists as a source of information about climate change, vaccines, and forest management, while those that identified with the Tea Party where the least likely to trust scientists in each policy domain [@hamiltonTrustScientistsClimate2015]. Also of note, @gauchatPoliticalContextScience2015 found that religious conservatism (belief that the Bible is divinely inspired and the word of God), Authoritarian beliefs, and political conservatism were associated with a decreased likelihood of support for the use of science in policymaking.  
     
A debate exists about the degree to which the decline of trust in science among conservatives is due to the fact that conservatives and/or Republicans are particularly prone to be opposed to science [@conwayWhyConservativesTurned2012; @mooneyRepublicanBrainScience2012]. For example, some have argued that conservatism is associated with other characteristics such as a low tolerance for uncertainty, less openness to experience, intolerance of ambiguity, and less open-minded cognition, which may make conservatives less inclined towards science or scientific thinking [@jostPoliticalConservatismMotivated2003; @priceOpenMindedCognition2015].  However, @kahanNotePerverseEffects2016 found that an increasing propensity for actively open-minded thinking was not associated with agreement across political beliefs regarding climate change, but rather it was associated with increased polarization. This finding suggests that differences in open-minded cognition is not a sufficient explanation for differences across political beliefs with regard to climate change.^[As noted by @kahanNotePerverseEffects2016 (4) a study by @haranRoleActivelyOpenMinded2013 found that actively open-minded thinking is correlated with other factors (e.g., openness to experience) thought to be associated with conservatism that might also be associated with views about science.] 

Another explanation for declining trust in science among conservatives is that science and scientific evidence is often used to identify problems that then justify government action. In general, conservatives are less favorable to government regulation of economic activity and such regulation is often presented as a way to address problems identified by science. Since conservatives are distrustful of the solution, they become distrustful of the science that identified the problem, through a process termed _solution aversion_ [@campbellSolutionAversionRelation2014]. Along similar lines, other work has distinguished between _production_ scientists that work to develop technological advancements for industry and _impact_ scientists whose work is likely to identify health or environmental problems, and find that conservatives place greater trust in production scientists and less trust in impact scientists than liberals [@mccrightInfluencePoliticalIdeology2013]. Therefore, the views of conservatives regarding science is not an aversion to science per se, but rather to the solution -- government regulation -- to address the problems identified by scientists. 

A final explanation is that views of science are issue-based and there is no asymmetric polarization across political beliefs. Therefore, science is doubted when it leads to conclusions that go against an individual's values or beliefs [@washburnScienceDenialPolitical2018; @kraftWhyPeopleDon2015]. Noted examples are evolution for religious conservatives and climate change for economic conservatives; however, liberals and/or Democrats are just as susceptible to distrusting science when scientists are presented as favoring things with which go against liberal / Democratic political beliefs such as nuclear power or fracking [@nisbetPartisanBrainHow2015; @dixonNeutralizingEffectPolitical2018a]. Additionally, who is viewed as an expert has been shown to be connected to whether or not the expert supports an individual's view of a particular issue, and this finding is consistent across political beliefs [@kahanCulturalCognitionScientific2011; @lachapellePublicPerceptionsExpert2014; @gauchatCulturalCognitiveMappingScientific2018; @suldovskyPublicPerceptionsWho2019].       

Given the connections between political beliefs and views about science, it is likely that political beliefs will influence beliefs about technocracy. Based on the observed declining trust of science among conservatives, I posit the following hypotheses: 

> H1: Liberal Democrats are _more_ likely to support technocratic decision-making than those with other types of political beliefs 

> H2: Conservative Republicans are _less_ likely to support technocratic decision-making than those with other types of political beliefs 

Going further, it is likely that views about technocracy will influence views about energy and environmental issues. 

# Technocracy, Energy, and Climate Change 

As noted, there is little work on how the public views technocracy and the implication of those views on other policy issues. However, as before there is a related literature on views about science and the use of science in policymaking. I draw on this literature to develop hypotheses about the influence of support for technocratic decision-making on support for energy and climate approaches.   

In general, research has shown support for the engagement of scientists in the policymaking process and that support is associated with views about several issues. For example, @nisbetUnderstandingPublicOpinion2014 examined the US public's views about science and society and support for stem cell research. Based on survey responses regarding science and society, respondents were placed in one of four categories including scientific optimists, scientific pessimists, conflicted, and disengaged. They found that scientific optimism was a stronger predictor of stem cell research support than partisanship, ideology, or religious views. Addition work found that trust in science and technology is associated with increased trust in specific technologies such as genetically modified organisms (GMOs), biotechnology, nuclear power, and wind power [@robertsCausalSpuriousRelationship2013]. Work by @huntRoleScientificSource2020, exploring scientific source credibility -- based on measures of scientist's understanding, integrity, agreement, concern, trust, and goodwill (lack of political bias; not career motivated) -- found that that the public's views on scientific credibility influenced their views on genetically modified food. Finally, when the public views scientists as knowledgeable, reflective of scientific consensus, and as serving the national interests they are more likely to support scientists making decisions over elected officials, with the most important predictor being that scientists served the national interest [@obrienScientificAuthorityPolicy2013].   
   
Based on the previous work regarding science and policymaking, I posit the following hypothesis: 

> H3: As support for technocratic decision-making increases so does support for energy sources and climate policies 

As discussed above, the influence of political beliefs on the role of science in policymaking is likely based on solution-aversion and/or the nature of the particular issue. Additionally, it is likely that views about technocratic versus democratic decision-making will be driven, in part, by some of the same factors that drive views about science in policymaking. Therefore, to examine support for technocracy and its impact on views about particular issues, I examine issues likely to invoke solution-aversion in conservative Republicans (climate change), yet be supported by liberal Democrats, as well as energy sources that are likely to be opposed by liberal Democrats (nuclear, pipelines, fracking, and offshore drilling), yet supported by conservative Republicans. In general, I expect liberal Democrats (conservative Republicans) to be less (more) supportive of all energy sources than those with other political beliefs and conservative Republicans (liberal Democrats) to less (more) supportive of all climate policies. However, I also expect there to be interaction effects between support for technocracy and political beliefs with regard to both energy sources and climate policies. 

Given that the nature of support for technocracy as well as views about particular energy and environmental issues are impacted by political beliefs, I expect that the influence of technocracy support on views about issues will not be uniform across political beliefs. A number of studies have found significant interaction effects between political beliefs and factors such as education, open-minded cognition, science knowledge, and numeracy with regard to issue-specific views, with the result of these interactions being increased polarization [@drummondIndividualsGreaterScience2017; @kahanPolarizingImpactScience2012; @joslynWhoKnowsBest2014a; @kahanNotePerverseEffects2016; @hamiltonEducationPoliticsOpinions2011]. Therefore, it's possible that increasing support for technocratic decision-making might also increase polarization with regard to energy and environmental issues. However, it is also possible that technocracy support would decrease polarization. The mechanism that is assumed to lead to polarization with increased education or science knowledge is that those that are more sophisticated are better "motivated reasoners" and are better able to connect views on issues with underlying political beliefs. It is not likely that support for technocracy would operate the same way because support for technocracy implies giving decision-making authority to experts rather than political actors. As a result, increasing support for technocracy would have a "pull" effect and likely be more impactful for those less likely to support controversial energy sources or climate policies. Therefore, I posit:      

> H4: As support for technocratic decision-making increases so does support for energy sources and climate policies, but not consistently across all types of political beliefs 

The expected inconsistency of the impact of technocracy support across political beliefs leaves open the possibility of polarization or convergence of views about energy sources and climate policies as support for technocracy increases. In the next section, I discuss the data and variables used to test the hypotheses. 

# Data and Variables
```{r setup, include=FALSE, message=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r models, include=FALSE, message=FALSE}
# load data
library("tidyverse")
source("scripts/recodes.R")

# science/policymaking model 
natData$sciPolScale3 <- round((natData$sciPol1+natData$sciPol2+natData$sciPol3)/3,0)
sciPolicyA <- psy::cronbach(data.frame(natData$sciPol1,natData$sciPol2,natData$sciPol3))

summary(ols1<-lm(sciPolScale3~libDem+modConDem+modLibRep+conRep+evanProt+age+male+white+edu+inc, data=natData))

stargazer::stargazer(ols1, type="html", title="Table 1: OLS Estimates of the Technocratic Scale", align=TRUE,  dep.var.labels = c("Technocratic Scale"),covariate.labels = c("Liberal Democrats","Moderate/Conservative Democrats", "Moderate/Liberal Republicans","Conservative Republicans","Evangelical Protestants","Age","Gender (Male=1)","Race (White=1)","Education","Income"), omit.stat=c("f","rsq","ll","ser"), star.cutoffs = c(0.05, 0.01, 0.001),  no.space=TRUE, header=FALSE, out = "manuscript/table1.doc")

# energy models 
summary(ols2<-lm(nuclear~sciPolScale3+libDem+modConDem+modLibRep+conRep+evanProt+age+male+white+edu+inc, data=natData))
summary(ols2in<-lm(nuclear~sciPolScale3*libDem+sciPolScale3*modConDem+sciPolScale3*modLibRep+sciPolScale3*conRep+evanProt+
                     age+male+white+edu+inc, data=natData))

summary(ols3<-lm(pipelines~sciPolScale3+libDem+modConDem+modLibRep+conRep+evanProt+age+male+white+edu+inc, data=natData))
summary(ols3in<-lm(pipelines~sciPolScale3*libDem+sciPolScale3*modConDem+sciPolScale3*modLibRep+sciPolScale3*conRep+evanProt+
                     age+male+white+edu+inc, data=natData))

summary(ols4<-lm(fracking~sciPolScale3+libDem+modConDem+modLibRep+conRep+evanProt+age+male+white+edu+inc, data=natData))
summary(ols4in<-lm(fracking~sciPolScale3*libDem+sciPolScale3*modConDem+sciPolScale3*modLibRep+sciPolScale3*conRep+evanProt+
                     age+male+white+edu+inc, data=natData))

summary(ols5<-lm(offshoreDrill~sciPolScale3+libDem+modConDem+modLibRep+conRep+evanProt+age+male+white+edu+inc, data=natData))
summary(ols5in<-lm(offshoreDrill~sciPolScale3*libDem+sciPolScale3*modConDem+sciPolScale3*modLibRep+sciPolScale3*conRep+evanProt+
                     age+male+white+edu+inc, data=natData))

stargazer::stargazer(ols2,ols2in,ols3,ols3in,ols4,ols4in,ols5,ols5in, type="html", title="Table 2: OLS Estimates of Energy Preferences", align=TRUE, dep.var.labels = c("Nuclear", "Pipelines", "Fracking", "Offshore Drilling"), column.labels = c("Main","Interaction","Main","Interaction","Main","Interaction","Main","Interaction"), covariate.labels = c("Technocratic Scale","Liberal Democrats","Moderate/Conservative Democrats", "Moderate/Liberal Republicans","Conservative Republicans","Evangelical Protestants","Age","Gender (Male=1)","Race (White=1)","Education","Income", "Tech*Lib Dem", "Tech*Mod/Con Dem", "Tech*Mod/Lib Rep", "Tech*Con Rep"), omit.stat=c("f","rsq","ll","ser"),star.cutoffs = c(0.05, 0.01, 0.001), float.env = "sidewaystable", font.size="footnotesize", label=c("tab:enModels"), no.space=TRUE, header=FALSE, out = "manuscript/table2.doc")

# climate models 
summary(ols6<-lm(gccIntagree~sciPolScale3+libDem+modConDem+modLibRep+conRep+evanProt+age+male+white+edu+inc, data=natData))
summary(ols6in<-lm(gccIntagree~sciPolScale3*libDem+sciPolScale3*modConDem+sciPolScale3*modLibRep+sciPolScale3*conRep+evanProt+
                     age+male+white+edu+inc, data=natData))

summary(ols7<-lm(gccEpa~sciPolScale3+libDem+modConDem+modLibRep+conRep+evanProt+age+male+white+edu+inc, data=natData))
summary(ols7in<-lm(gccEpa~sciPolScale3*libDem+sciPolScale3*modConDem+sciPolScale3*modLibRep+sciPolScale3*conRep+evanProt+
                     age+male+white+edu+inc, data=natData))

summary(ols8<-lm(gccTax~sciPolScale3+libDem+modConDem+modLibRep+conRep+evanProt+age+male+white+edu+inc, data=natData))
summary(ols8in<-lm(gccTax~sciPolScale3*libDem+sciPolScale3*modConDem+sciPolScale3*modLibRep+sciPolScale3*conRep+evanProt+
                     age+male+white+edu+inc, data=natData))

summary(ols9<-lm(gccCapTrade~sciPolScale3+libDem+modConDem+modLibRep+conRep+evanProt+age+male+white+edu+inc, data=natData))
summary(ols9in<-lm(gccCapTrade~sciPolScale3*libDem+sciPolScale3*modConDem+sciPolScale3*modLibRep+sciPolScale3*conRep+evanProt+
                     age+male+white+edu+inc, data=natData))

stargazer::stargazer(ols6,ols6in,ols7,ols7in,ols8,ols8in,ols9,ols9in, type="html", title="Table 3: OLS Estimates of Climate Policy Preferences", align=TRUE, dep.var.labels = c("International Agreement", "EPA Regulations", "Carbon Tax", "Cap-and-Trade"), column.labels = c("Main","Interaction","Main","Interaction","Main","Interaction","Main","Interaction"), covariate.labels = c("Technocratic Scale","Liberal Democrats","Moderate/Conservative Democrats", "Moderate/Liberal Republicans","Conservative Republicans","Evangelical Protestants","Age","Gender (Male=1)","Race (White=1)","Education","Income", "Tech*Lib Dem", "Tech*Mod/Con Dem", "Tech*Mod/Lib Rep", "Tech*Con Rep"),omit.stat=c("f","rsq","ll","ser"),star.cutoffs = c(0.05, 0.01, 0.001), float.env = "sidewaystable", font.size="footnotesize", label=c("tab:gccModels"), no.space=TRUE, header=FALSE, out = "manuscript/table3.doc")

corPB <- cor.test(natData$ideology, natData$partisan)

```

```{r graphs, include=FALSE, message=FALSE}
## graphs
libDemData <- data.frame(
  sciPolScale3=1:7,
  libDem=1,
  modConDem=mean(natData$modConDem, na.rm=TRUE),
  modLibRep=mean(natData$modLibRep, na.rm=TRUE),
  conRep=mean(natData$conRep, na.rm=TRUE),
  evanProt=mean(natData$evanProt, na.rm=TRUE),
  age=mean(natData$age, na.rm=TRUE),
  male=mean(natData$male, na.rm=TRUE),
  white=mean(natData$white, na.rm=TRUE),
  edu=mean(natData$edu, na.rm=TRUE),
  inc=mean(natData$egal, na.rm=TRUE)
)


modConDemData <- data.frame(
  sciPolScale3=1:7,
  libDem=mean(natData$libDem, na.rm=TRUE),
  modConDem=1,
  modLibRep=mean(natData$modLibRep, na.rm=TRUE),
  conRep=mean(natData$conRep, na.rm=TRUE),
  evanProt=mean(natData$evanProt, na.rm=TRUE),
  age=mean(natData$age, na.rm=TRUE),
  male=mean(natData$male, na.rm=TRUE),
  white=mean(natData$white, na.rm=TRUE),
  edu=mean(natData$edu, na.rm=TRUE),
  inc=mean(natData$egal, na.rm=TRUE)
)


modLibRepData <- data.frame(
  sciPolScale3=1:7,
  libDem=mean(natData$libDem, na.rm=TRUE),
  modConDem=mean(natData$modConDem, na.rm=TRUE),
  modLibRep=1,
  conRep=mean(natData$conRep, na.rm=TRUE),
  evanProt=mean(natData$evanProt, na.rm=TRUE),
  age=mean(natData$age, na.rm=TRUE),
  male=mean(natData$male, na.rm=TRUE),
  white=mean(natData$white, na.rm=TRUE),
  edu=mean(natData$edu, na.rm=TRUE),
  inc=mean(natData$egal, na.rm=TRUE)
)


conRepData <- data.frame(
  sciPolScale3=1:7,
  libDem=mean(natData$libDem, na.rm=TRUE),
  modConDem=mean(natData$modConDem, na.rm=TRUE),
  modLibRep=mean(natData$modLibRep, na.rm=TRUE),
  conRep=1,
  evanProt=mean(natData$evanProt, na.rm=TRUE),
  age=mean(natData$age, na.rm=TRUE),
  male=mean(natData$male, na.rm=TRUE),
  white=mean(natData$white, na.rm=TRUE),
  edu=mean(natData$edu, na.rm=TRUE),
  inc=mean(natData$egal, na.rm=TRUE)
)


libDemNuke <- predict(ols2in, libDemData, se.fit=TRUE)
modConDemNuke <- predict(ols2in, modConDemData, se.fit=TRUE)
modLibRepNuke <- predict(ols2in, modLibRepData, se.fit=TRUE)
conRepNuke <- predict(ols2in, conRepData, se.fit=TRUE)


libDemPipe <- predict(ols3in, libDemData, se.fit=TRUE)
modConDemPipe <- predict(ols3in, modConDemData, se.fit=TRUE)
modLibRepPipe <- predict(ols3in, modLibRepData, se.fit=TRUE)
conRepPipe <- predict(ols3in, conRepData, se.fit=TRUE)

libDemFrack <- predict(ols4in, libDemData, se.fit=TRUE)
modConDemFrack <- predict(ols4in, modConDemData, se.fit=TRUE)
modLibRepFrack <- predict(ols4in, modLibRepData, se.fit=TRUE)
conRepFrack <- predict(ols4in, conRepData, se.fit=TRUE)

libDemDrill <- predict(ols5in, libDemData, se.fit=TRUE)
modConDemDrill <- predict(ols5in, modConDemData, se.fit=TRUE)
modLibRepDrill <- predict(ols5in, modLibRepData, se.fit=TRUE)
conRepDrill <- predict(ols5in, conRepData, se.fit=TRUE)

# postscript("fig-output/fig1.eps", horizontal = FALSE, onefile = FALSE, paper = "special", width = 10, height = 12)
png("fig-output/fig1.png", width = 480, height = 520)
par(oma = c(1.5,1,1.5,1), mfrow=c(2,2))
plot(libDemNuke$fit, type="l", ylim=c(0.3,7), lwd=3, ylab="", xlab="", main="Nuclear")
lines(modConDemNuke$fit, lty=2, lwd=3)
lines(modLibRepNuke$fit, lty=3, lwd=3)
lines(conRepNuke$fit, lty=4, lwd=3)
# legend("topleft", c("Liberal Democrats","Conservative Republican"), lty=1:4, bty="n", cex=1)

plot(libDemPipe$fit, type="l", ylim=c(1,7), lwd=3, ylab="", xlab="", main="Pipelines")
lines(modConDemPipe$fit, lty=2, lwd=3)
lines(modLibRepPipe$fit, lty=3, lwd=3)
lines(conRepPipe$fit, lty=4, lwd=3)
#legend("topleft", c("Hierarchs","Egalitarians","Indiviualists", "Fatalist"), lty=1:4, bty="n", cex=1.5)

plot(libDemFrack$fit, type="l", ylim=c(1,7), lwd=3, ylab="", xlab="", main="Fracking")
lines(modConDemFrack$fit, lty=2, lwd=3)
lines(modLibRepFrack$fit, lty=3, lwd=3)
lines(conRepFrack$fit, lty=4, lwd=3)
#legend("topleft", c("Hierarchs","Egalitarians","Indiviualists", "Fatalist"), lty=1:4, bty="n", cex=1.5)

plot(libDemDrill$fit, type="l", ylim=c(1,7), lwd=3, ylab="", xlab="", main="Offshore Drilling")
lines(modConDemDrill$fit, lty=2, lwd=3)
lines(modLibRepDrill$fit, lty=3, lwd=3)
lines(conRepDrill$fit, lty=4, lwd=3)
#legend("topleft", c("Hierarchs","Egalitarians","Indiviualists", "Fatalist"), lty=1:4, bty="n", cex=1.5)
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("topleft", c("Liberal Democrats","Moderate/Conservative Democrats"), lwd=4, lty=1:2,
       xpd=TRUE, horiz=TRUE, bty="n",cex=1)
legend("topright", c("Moderate/Liberal Republicans","Conservative Republicans"), lwd=4, lty=3:4,
       xpd=TRUE, horiz=TRUE, bty="n",cex=1)
legend("bottom", c("Technocratic Scale"),xpd=TRUE, horiz=TRUE, bty="n",cex=1.75)
dev.off()

# climate policy 
libDemIntAgree <- predict(ols6in, libDemData, se.fit=TRUE)
modConDemIntAgree <- predict(ols6in, modConDemData, se.fit=TRUE)
modLibRepIntAgree <- predict(ols6in, modLibRepData, se.fit=TRUE)
conRepIntAgree <- predict(ols6in, conRepData, se.fit=TRUE)

libDemEPA <- predict(ols7in, libDemData, se.fit=TRUE)
modConDemEPA <- predict(ols7in, modConDemData, se.fit=TRUE)
modLibRepEPA <- predict(ols7in, modLibRepData, se.fit=TRUE)
conRepEPA <- predict(ols7in, conRepData, se.fit=TRUE)

libDemTax <- predict(ols8in, libDemData, se.fit=TRUE)
modConDemTax <- predict(ols8in, modConDemData, se.fit=TRUE)
modLibRepTax <- predict(ols8in, modLibRepData, se.fit=TRUE)
conRepTax <- predict(ols8in, conRepData, se.fit=TRUE)

libDemCapTrade <- predict(ols9in, libDemData, se.fit=TRUE)
modConDemCapTrade <- predict(ols9in, modConDemData, se.fit=TRUE)
modLibRepCapTrade <- predict(ols9in, modLibRepData, se.fit=TRUE)
conRepCapTrade <- predict(ols9in, conRepData, se.fit=TRUE)

# postscript("fig-output/fig2.eps", horizontal = FALSE, onefile = FALSE, paper = "special", width = 10, height = 12)
png("fig-output/fig2.png", width = 480, height = 520)
par(oma = c(1.5,1,1.5,1), mfrow=c(2,2))
plot(libDemIntAgree$fit, type="l", ylim=c(1,7), lwd=3, ylab="", xlab="", main="International Agreement")
lines(modConDemIntAgree$fit, lty=2, lwd=3)
lines(modLibRepIntAgree$fit, lty=3, lwd=3)
lines(conRepIntAgree$fit, lty=4, lwd=3)
#legend("topleft", c("Hierarchs","Egalitarians","Indiviualists", "Fatalist"), lty=1:4, bty="n", cex=1.5)

plot(libDemEPA$fit, type="l", ylim=c(1,7), lwd=3, ylab="", xlab="", main="EPA Regulations")
lines(modConDemEPA$fit, lty=2, lwd=3)
lines(modLibRepEPA$fit, lty=3, lwd=3)
lines(conRepEPA$fit, lty=4, lwd=3)
#legend("topleft", c("Hierarchs","Egalitarians","Indiviualists", "Fatalist"), lty=1:4, bty="n", cex=1.5)

plot(libDemTax$fit, type="l", ylim=c(1,7), lwd=3, ylab="", xlab="", main="Carbon Tax")
lines(modConDemTax$fit, lty=2, lwd=3)
lines(modLibRepTax$fit, lty=3, lwd=3)
lines(conRepTax$fit, lty=4, lwd=3)
#legend("topleft", c("Hierarchs","Egalitarians","Indiviualists", "Fatalist"), lty=1:4, bty="n", cex=1.5)

plot(libDemCapTrade$fit, type="l", ylim=c(1,7), lwd=3, ylab="", xlab="", main="Cap-and-Trade")
lines(modConDemCapTrade$fit, lty=2, lwd=3)
lines(modLibRepCapTrade$fit, lty=3, lwd=3)
lines(conRepCapTrade$fit, lty=4, lwd=3)
#legend("topleft", c("Hierarchs","Egalitarians","Indiviualists", "Fatalist"), lty=1:4, bty="n", cex=1.5)
par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
legend("topleft", c("Liberal Democrats","Moderate/Conservative Democrats"), lwd=4, lty=1:2,
       xpd=TRUE, horiz=TRUE, bty="n",cex=1)
legend("topright", c("Moderate/Liberal Republicans","Conservative Republicans"), lwd=4, lty=3:4,
       xpd=TRUE, horiz=TRUE, bty="n",cex=1)
legend("bottom", c("Technocratic Scale"),xpd=TRUE, horiz=TRUE, bty="n",cex=1.75)
dev.off()

```


To test the hypotheses discussed above, I use original survey data collected through a sample of the US population (18+) obtained by Survey Sample International (SSI) with US Census based quotas for age, gender, and ethnicity. The data includes about 1200 respondents and was collected in October of 2017. The survey asked a battery of questions regarding political beliefs, views of science in policymaking, energy preferences, climate policy views, as well as demographics.^[More information about the survey respondents and the questions are in the Appendix.] 

To measure views about technocracy, I draw on three statements about the role of expertise and knowledge versus the public with regard to decision-making for energy and environmental issues.^[The technocracy questions used here are similar to those used by @silvaReconcilingScientistsBeliefs2007.] Respondents were asked to rate their agreement with each statement on a 1 to 7 scale, with 1 = strongly disagree and 7 = strongly agree. The questions and their mean values on the 1 to 7 scale are:  

* _Scientists and technically trained experts, not the public, should make decisions about the applications of advanced technologies to address energy needs and environmental concerns_: Mean = `r round(mean(natData$sciPol1,na.rm=T),2)`  

* _Those who are better informed and knowledgeable should have greater influence in policy making_: Mean = `r round(mean(natData$sciPol2,na.rm=T),2)`   

* _Technical issues are so complex that most people cannot contribute to reasonable policy choices_: Mean = `r round(mean(natData$sciPol3,na.rm=T),2)`   

The questions were combined and averaged to create a _technocratic scale_ with a mean value of `r round(mean(natData$sciPolScale3,na.rm=T),3)` on the 1 to 7 scale, indicating a relative propensity for support of technocratic decision-making with regard to energy and environmental concerns. Additionally, Cronbach's $\alpha$ = `r round(sciPolicyA$alpha,2)`, which indicates acceptable scale reliability. 

Respondents were asked their views on several controversial energy sources including nuclear; pipelines for oil and gas; hydraulic fracturing (fracking); and offshore drilling for oil and natural gas. Specifically, respondents were asked their support for each energy source on a 1 = strongly oppose to 7 = strongly support scale, and the average level of support for each energy source were _nuclear_ = `r round(mean(natData$nuclear,na.rm=T),2)`, _pipelines_ = 
`r round(mean(natData$pipelines,na.rm=T),2)`, _fracking_ = `r round(mean(natData$fracking,na.rm=T),2)` and _offshore drilling_ = 
`r round(mean(natData$offshoreDrill,na.rm=T),2)`. 

With regard to climate policy, respondents were asked their support on a 1 = strongly oppose to 7 = strongly support scale for the following climate policies: _international agreement_ = `r round(mean(natData$gccIntagree,na.rm=T),2)`, _Environmental Protection Agency (EPA) regulations_ =  `r round(mean(natData$gccEpa,na.rm=T),2)`, _carbon tax_ = `r round(mean(natData$gccTax,na.rm=T),2)`, and _cap-and-trade_ =  `r round(mean(natData$gccCapTrade,na.rm=T),2)`.^[Exact question wording for the energy and climate questions is available in the Appendix. Of note, the energy questions begin with a prompt that states the following are actions that private businesses and/or state and local governments can take to resolve some community concerns, whereas, the climate change questions are presented as policy options for the federal government. It is possible that the private businesses versus federal government prompts of the energy and climate questions could exacerbate political differences. However, as noted these issues were chosen because of their polarized nature.] 

For political beliefs, I use a combined measure of ideology and partisanship. To measure ideology respondents were asked to place themselves on a 7-point scale where 1 = strongly liberal, 4 = middle of the road, and 7 = strongly conservative. Those that placed themselves a 1, 2, or 3 on the ideology scale were coded as liberal, those that placed themselves as a 5, 6, or 7 were coded as conservative, and those that placed themselves at a 4 were coded as moderate. Partisanship was measured by asking respondents with which party they identified including Democratic, Republican, independent, other/no party. Those that answered independent or no/party were asked to which party they leaned, Democratic or Republican and they were subsequently assigned to the party toward which they leaned. The ideology and partisanship measures were combined and dummy variables were created for liberal Democrats, moderate/conservative Democrats, moderate/liberal Republicans, and conservative Republicans.^[The seven-point ideology and partisanship scales had a correlation coefficient of `r round(corPB$estimate, 3)`. Additionally, combining ideology and party attachments allows for the inclusion of the cognition associated with ideology as well as the affective elements associated with political party attachments. Also, similar measures have been used in other studies [e.g., @kahanNotePerverseEffects2016 and @gustafsonDevelopmentPartisanPolarization2019].] Moderates and true independents are the excluded referent in the subsequent analysis. 

I also control for religious beliefs using a dummy variable for Evangelical Protestants, which have been the most skeptical of science in general [@ecklundExaminingLinksReligion2016; @ellisonConservativeProtestantismPublic1995], and climate change, specifically [@smithAmericanEvangelicalsGlobal2013]. Finally, I control for demographics including age, gender, race/ethnicity, education, and income. 

# Analysis and Results 

The first analysis uses OLS regression to examine H1 and H2, which posited that support for technocracy was more likely among liberal Democrats (H1) and less likely among conservative Republicans (H2) than other political beliefs. As noted, support for technocracy is based on the 1 to 7 technocratic scale, political beliefs are a set of dummy variables with moderate independent as the referent group, and the set of controls include a dummy variable for Evangelical Protestants as well as several demographic variables. The results are shown in Table 1.  

**Table 1 here** 

As shown and as expected, liberal Democrats were more likely than moderate independents to express support for technocratic decision-making regarding energy and environmental issues. However, against expectations conservative Republicans were no more or less likely to support technocracy than were moderate independents. Additionally, both moderate to conservative Democrats and moderate to liberal Republicans were more likely than moderate independents to support technocracy. Evangelical Protestants were no more of less likely than those of other religious affiliations to support technocratic decision-making. Finally, older respondents were less likely than younger respondents to support technocracy, whereas, males, whites, and those with more education were more supportive than females, non-white, and less educated respondents. 

Next, I use OLS regression to examine support for the various types of controversial energy sources based on views about technocracy, political beliefs, and demographics. I expect that as support for technocracy increases, so does support for each energy source (H3). In addition, I also examine the interaction between the technocratic scale and each type of political belief and I expect the impact of technocracy support on support of energy sources to vary across political beliefs. The results are shown in Table 2. 

**Table 2 here**

Looking first to the main effects (columns 1,3,5, and 7 in Table 2), as expected an increase in support for technocracy is associated with an increase in support for all types of energy. Also as expected, liberal Democrats are _less_ supportive and conservative Republicans are _more_ supportive of all energy sources than moderate independents, all else equal. For the other political beliefs moderate/conservative Democrats and moderate/liberal Republicans are _more_ supportive than moderate independents of all types but nuclear energy. In terms of the control variables, being an Evangelical Protestant is associated with an increase in support for all types of energy as is being male as opposed to female. Education is associated with less support for all types except for nuclear, whereas, increasing income is associated with increasing support for all types of energy but nuclear. 

In terms of the interaction terms (columns 2,4,6, and 8 in Table 2), there are positive and significant interactions for liberal Democrats and the technocratic scale for fracking and offshore drilling. For moderate/conservative Democrats, the interactions were significant for all types of energy and for moderate/liberal Republicans all interactions were significant expect for nuclear energy. Finally, no interactions were significant for conservative Republicans. The interaction results are graphed in Figure 1. 

**FIGURE 1 ABOUT HERE: Figure 1: Support for Energy Sources by Political Beliefs and Technocratic Views** 

As shown in Figure 1, as the technocratic scale increases so does support for each energy source across all types of political beliefs; however, those increases are not uniform. Of note, those political beliefs more in the middle -- moderate/conservative Democrats and moderate/liberal Republicans -- seem to be those most responsive to increases in the technocratic scale. Also of note, liberal Democrats and conservative Republicans maintain a roughly 2 point difference on the seven-point support scale for pipelines, fracking, and offshore drilling across all values of the technocratic scale even with a modest interaction effect for liberal Democrats regarding fracking and offshore drilling. Finally, all political beliefs except for liberal Democrats seem to converge on support for all energy types as they reach the maximum level of support for technocratic decision-making.       

For the climate policies, I expect that increasing support for technocratic decision-making to be associated with increasing support for all climate policies, liberal Democrats (conservative Republicans) to be more (less) supportive of all climate policies than moderate independents, and interaction effects between political beliefs and support for technocracy. The results are shown in Table 3 and, as before, I first examine the main effect shown in columns 1, 3, 5, and 7 and then the interactions shown in columns 2, 4, 6, and 8. 

**Table 3 here** 

As can be seen and as expected, as respondents become more supportive of technocratic decision-making they also become more supportive of each of the climate policies. In addition, liberal Democrats are more supportive of each climate policy, whereas, conservative Republicans are less supportive of each policy than moderate independents. The other types of political beliefs show no significant differences from moderate independents in support of any of the climate policies. In terms of demographics, older respondents are less supportive of a cap-and-trade policy than are younger respondents and male respondents are less likely than female respondents to support an international agreement or EPA regulations to address climate change. Finally, more educated respondents are more supportive of carbon pricing policies -- a carbon tax and cap-and-trade -- than are less educated respondents and there are no differences with regard to Evangelical Protestants or income. 

For the interactions shown in columns 2, 4, 6, and 8 of Table 3 as well as Figure 2, liberal Democrats become slightly less supportive of international agreements, EPA regulations, and a carbon tax as they become more supportive of technocratic decision-making. As shown in Figure 2, this negative interaction indicates that increases in the technocratic scale matter less for climate policy support for liberal Democrats than it does for moderate independents. In other words, liberal Democrats are likely to have a baseline of strong support for climate policy, therefore, their views on technocracy have less of an impact on their climate policy views than for other types of political beliefs. For conservative Republicans, the interactions are positive and significant indicating that increasing support for technocracy is more impactful for their support of each climate policy than it is for moderate independents. In terms of the other types of political beliefs, only the interaction between the technocratic scale and moderate/conservative Democrats with views on cap-and-trade is statistically significant. Finally, as seen in Figure 2 all types of political beliefs converge on support for each climate policy at the highest level of support for technocracy.     

**FIGURE 2 ABOUT HERE: Figure 2: Support for Climate Policies by Political Beliefs and Technocratic Views** 

# Discussion 

Little work to date has explicitly examined public support for technocracy (vs democracy) and the implication of that support for views about controversial energy and climate policies. In this paper, I employed a seven-point technocratic scale based on responses to three questions. Drawing on literature regarding views about science, I developed and tested hypotheses regarding political beliefs and support for technocratic decision-making. Overall, I found that liberal Democrats, moderate/conservative Democrats, and moderate/liberal Republicans were more likely than moderate independents to support technocratic decision-making. However, contrary to expectations conservatives Republicans were no less likely than moderate independents to support technocracy. It may be the case that conservative Republican are not opposed to expert-driven decision-making per se, rather other types of political beliefs are just more supportive [see @blankDoesPartisanshipShape2015 for a similar finding regarding scientific advice in policymaking]. These results indicate that in polarized issues like energy and climate, political beliefs are an important factor for not only determining what policies should be adopted but also for determining who it is that ought to be making decisions in complex policy areas.

Going beyond views about technocracy, I also examined the influence of those views on specific, controversial issues. I explored energy sources expected to garner opposition from liberal Democrats as well as climate policies expected to be opposed by conservative Republicans.  As expected, I found that, on average, an increase in support for technocracy was associated with increases in support for all energy sources and climate policies. These findings suggest that support for expert-driven decision-making is associated with support for approaches that require expert knowledge to develop, implement, and monitor.  

Also as expected, the influence of technocracy support was not uniform across all types of political beliefs. In general, there was a trend of increasing support for energy sources and climate policies as support for technocracy increased across all types of political beliefs; however, there were some significant interactions effects. As discussed above, increasing support for technocratic decision-making could have been similar to measures of sophistication (e.g., education, science knowledge) and lead to an increase in polarization across political beliefs or, because technocracy is premised on taking authority away from political actors, increasing technocracy might conceivably lead to less polarization. Overall, the findings point to the latter. With regard to energy sources, the more moderate beliefs -- Moderate/Conservative Democrats, and Moderate/Liberal Republicans -- were the most impacted by increasing support for technocratic decision-making, whereas conservative Republicans and liberal Democrats showed less of an interactive effect. Yet, support increased for all energy sources as support for technocracy increased across all political beliefs. For climate policies, it was with liberal Democrats and conservative Republicans where significant interactions occurred. I had conjectured that support for technocracy could have a pull effect with the largest impact being on those who were less supportive of the energy sources or climate policies based on political beliefs; however, that seems to have only been the case with conservative Republicans and climate policy  (see Figure 2). For liberal Democrats, increasing technocracy support did not seem to have any large interactive effects with regard to support for energy sources. Finally, apart from liberal Democrats and energy sources, there tended to be a convergence of views, rather than increased polarization, about energy and climate policies across political beliefs at the highest level of support for technocracy. This finding suggests that views about technocracy may be an important factor in guiding the US public's views on complex energy and environmental issues.  


# Conclusion 

Technocratic decision-making, or technocracy, is an important feature of energy and environmental policymaking, and as energy and environmental issues become more complex and with growing technological capabilities, democratic societies will need to address the balance between technocratic and representative decision-making. Yet, little work has examined how the public views technocratic decision-making and the implications of those views for opinions about particular issues. In this paper, I explore public opinion regarding technocracy and how views about technocracy inform views about energy sources and climate policies. The findings suggest that political beliefs (ideology and partisanship) are a significant factor for predicting views about technocratic decision-making. Further, the findings suggest significant interactions effects between support for technocracy and political beliefs in the ways that they inform energy and climate views. Overall, these findings complement and extend findings regarding the role of science in policymaking and point to the need to better understand how the public weighs technocratic versus democratic decision-making.      

Future research should further explore technocracy support and its importance beyond energy and environmental issues. Additionally, future work should examine issues that are less polarized. The importance of political beliefs are likely to diminish when less polarized issues are considered. Future work should also directly compare views about technocracy with measures used previously to examine views about science in policymaking. While views about technocracy, particularly in comparison to more representative and democratic decision-making, are distinct from views about science it's likely that some of the factors such as trust, credibility, and perceived consensus among experts would influence both technocracy support and science in policymaking support. It is important to determine how technocracy views differ, or don't, from views about science and science in policymaking more broadly. Similarly, the connections between technocracy views and growing support for populism, with its associated anti-intellectualism, should also be explored.  

\newpage

# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\singlespace
\noindent
